Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2023-10-26 doc/manuscript/ForC-EFDB manuscript.Rmd
update refs

New data collection and analysis is needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [Table `r table_variables`, @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain very poorly quantified [Table `r table_variables`, @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labriere_forest_2023].

@@

Filling knowledge gaps in other EFDB-relevant variables will require more effort, but this effort is warranted given their importance for estimating forest C stock chnages.
Although aboveground biomass is the most studied variable considered here (Table `r table_variables`) and is the target of satellite missions [@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], significant ground-based research effort is required to create accurate global maps of forest biomass and changes therein [@duncanson_importance_2019; @labriere_forest_2023;@calders_laser_2022].
Given observations of increasing tree mortality in some forested regions [@mcdowell_pervasive_2020], better characterization of forest dead wood will be critical.
Additionally, C stocks in forest organic horizons and soils can be quite substantial and highly uncertain in many parts of the world [@tifafi_large_2018].
Significant investment in ground-based forest research will be critical to filling these gaps.

@@

We recommend that, in order to make research valuable to estimate C stock changes according to methods provided in the IPCC guidelines, researchers calculate and report results according to IPCC good practice (Table `r table_recommendations`).
It is particularly noteworthy that simple decisions on the presentation of results will determine whether the records meet the criteria for inclusion in EFDB.
Some examples are as follows:
(1) presenting data only in a figure makes them ineligible for inclusion in EFDB, whereas presentation in a table or supplementary data file allows inclusion while supporting FAIR (http://dx.doi.org/10.1038/d41586-019-01720-7) goals;
(2) direct presentation of all relevant variables allows inclusion, whereas presenting only components of variables of interest (e.g., parsing litter into fine woody debris, OL, OF, and OH layers) or requiring simple mathematical operations to obtain a variable of interest (e.g., *delta.agb* = *ANPP_woody* - *woody.mortality.agb*) disqualifies records from inclusion;
(3) matching IPCC-defined thresholds for defining C pools (Table `r table_pools`), which may vary by country, can make the data far more relevant estimating forest C stock changes according to IPCC guidelines (e.g., using a 10 cm cutoff between dead wood and litter, presenting soil C to a depth of 30 cm).
It should also be emphasized that reporting of 95% confidence intervals (or other metrics of error), when applicable, is highly desirable and makes the data more relevant to IPCC.

>>>>

New data collection and analysis is needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [Table `r table_variables`, @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain very poorly quantified [Table `r table_variables`, @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labrière_forest_2023].

@@

Filling knowledge gaps in other EFDB-relevant variables will require more effort, but this effort is warranted given their importance for estimating forest C stock chnages.
Although aboveground biomass is the most studied variable considered here (Table `r table_variables`) and is the target of satellite missions [@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], significant ground-based research effort is required to create accurate global maps of forest biomass and changes therein [@duncanson_importance_2019; @labrière_forest_2023;@calders_laser_2022].
Given observations of increasing tree mortality in some forested regions [@mcdowell_pervasive_2020], better characterization of forest dead wood will be critical.
Additionally, C stocks in forest organic horizons and soils can be quite substantial and highly uncertain in many parts of the world [@tifafi_large_2018].
Significant investment in ground-based forest research will be critical to filling these gaps.

@@

We recommend that, in order to make research valuable to estimate C stock changes according to methods provided in the IPCC guidelines, researchers calculate and report results according to IPCC good practice (Table `r table_recommendations`).
It is particularly noteworthy that simple decisions on the presentation of results will determine whether the records meet the criteria for inclusion in EFDB.
Some examples are as follows:
(1) presenting data only in a figure makes them ineligible for inclusion in EFDB, whereas presentation in a table or supplementary data file allows inclusion while supporting FAIR goals [@stall_make_2019];
(2) direct presentation of all relevant variables allows inclusion, whereas presenting only components of variables of interest (e.g., parsing litter into fine woody debris, OL, OF, and OH layers) or requiring simple mathematical operations to obtain a variable of interest (e.g., *delta.agb* = *ANPP_woody* - *woody.mortality.agb*) disqualifies records from inclusion;
(3) matching IPCC-defined thresholds for defining C pools (Table `r table_pools`), which may vary by country, can make the data far more relevant estimating forest C stock changes according to IPCC guidelines (e.g., using a 10 cm cutoff between dead wood and litter, presenting soil C to a depth of 30 cm).
It should also be emphasized that reporting of 95% confidence intervals (or other metrics of error), when applicable, is highly desirable and makes the data more relevant to IPCC.
